---
title: "Time Series Decomposition Based on Markov Random Fields"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document: default
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The full instructions for the research project are provided [here](https://uw-csss-564.github.io/2018/pages/research_project/). This file is a template to help you organize it. Delete any instructional text.

Succinctly describe the computational narrative.

# Setup

Load packages you will use the analysis.
Do not install any package here; it is impolite to install packages or change anything outside the project directory without the user's consent.
```{r setup,message=FALSE}
library("tidyverse")
```

You should also load data, define options or constant variables, or define utility functions in this section.

R markdown will always use the directory of the file as the working directory when rendering. 
This can be annoying since if the R markdown document is in a subdirectory, this will be different than the project working directory. I suggest using the **here** package for always constructing file paths relative to the root directory. 
You can read more about this [here](https://github.com/jennybc/here_here).
```{r paths}
data_dir <- here::here("data")
results_dir <- here::here("results")
src_dir <- here::here("src")
```


# Rest of the paper

Organize the rest of the computational narrative as appropriate.

Use sections to organize the computational narrative, similar to how you would use them in a research article.
Do not use sections merely to make text bold.

There should be no extraneous messages, warnings, or errors in the output. Use `message = FALSE` to suppress printing  message output, and resolve all warnings and errors in your code.

# Original Computational Environment

```{r}
sessionInfo()
```
